---
title: "Test1 Review"
author: "Sheila Weaver"
date: "9/14/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

```

### Review the ideas outlined in the Reading for chapters 1 and 2.

### Review these ideas from class:

* Data Structures:  vectors, data frames are the two we've used so far. (Others include matrices, arrays, lists)

* Data Types in R:  Logical, Numeric, Character/String

  + Quantitative vs Categorical data:  
  + Quant:   age, earnings
  + Categ:   year in school, response (yes/no)

* R is case sensitive

* How are missing values represented in R?  

* How do you get out of the + prompt?  

* Functions and arguments (which are which below?):
  + mean(Height)
  + sqrt(25)
  + mean(S$Earnings, na.rm=TRUE)


* What is the meaning of   <-   in R


* Programs and Packages:  R is a program;  dplyr, ggplot2, tidyverse are packages

* Installing and Loading:  How and when must you install/load packages?


###  Practice Writing Code

#####  Enter these vectors in R:

```{r entervectors}
#  Animal:  cat, dog, pig, opossum, horse, human
#  Sleep:  12.5, 10.1, 9.1, 18.0,  2.9, 8
#  REM:  3.2, 2.9, NA, 4.9, 0.6, 1.9   (don't put the NA in quotes)
#  Brain:  .026, .07, .18,.006, .655, 1.32
#  Body:  3.3, 14, 86.2, 1.7, 521, 62

```

##### Stats and new vectors

```{r vectorstats}
#  Have R calculate the mean Body size 

#  Have R calculate the mean REM sleep time

#  Do a summary of the Body vector

#  Create new vectors:  LBrain is the log of Brain and LBody is the log of Body
#  Note:  You may use the function log() OR the functionlog10()

```

##### Some plots

```{r plots}
#  Try some plots:   Brain by Body,  LBrain by LBody

#  Also:  Sleep by Body, then Sleep by LBody

```

##### Make and use a data frame

```{r dataframe}
#  Put the vectors in a data frame, called an:

#  Because it is small, you can print the data frame to make sure it looks correct.
an

#  Try these functions to examine the data frame further:  dim, head, tail, str, names.

#  Remove the vectors from your environment, using rm()

#  Do summary() on the whole data frame

#  Now, redo the plots using plot and an$

```





