---
title: "Bootstrap CI for a Mean"
author: "yourname"
date: "11/16/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

###  We want to estimate the true mean number of states visited by all students in this (and the other section of this) class.

#### We will start with a random sample of 5 students from the population (students in this class)

```{r}
#  Here is my original sample from the population
os <- c(13, 12, 25, 17, 5)

#  Find the PE.  Here the point estimate is the mean of the sample.
PE <- mean(os)
PE

#  To select a sample of 5 from the faked population
#  -- using our sample in place of the population:
#  Do two of these and calculate the mean from each
s <- sample(os, size = 5, replace = TRUE)
s
mean(s)


#  Record these means on your activity
```

### Loops

#### In order to develop code to have R make a bootstrap distribution, we need the coding idea of a LOOP.

```{r loops}
for (j in 1:5)  print("Hi there")

for (i in 1:5)  print(i)

for (i in 1:5) {  
  print("Hi")
  print(i)
  print(i^2)
}

```

### Saving the bootstrap means

```{r bootmeans}
# Now we want to save values in a vector, rather than just printing them on the console.  We'll call the vector Allmeans.

Allmeans <- c()
Allmeans
for (i in 1:10000) {
  s <- sample(os, size = length(os), replace = TRUE)
  Allmeans[i] <- mean(s)
}

#  Allmeans
hist(Allmeans)

SE <- sd(Allmeans)
SE

 # Note:  we used size = 5 above, to reflect the size of the original sample.  Using length(os) makes the code work for any length os vector.

# The values in Allmeans are the Statistics from bootstrap samples.  So the SE is just the standard deviation of Allmeans.

#  Make a histogram of Allmeans

```

###  Now, calculate the CI, using the same method as for percentages

```{r}
#  Calculate ME, lower and upper limits, then print in a sentence in terms of the problem.

ME <- 2 * SE

lowerlimit <- PE - ME
upperlimit <- PE + ME

lowerlimit <- round(lowerlimit, 2)
upperlimit <- round(upperlimit, 2)

print(paste("I'm 95% confident that the true mean number of states visited by students in 087 is captured in the interval from ", lowerlimit, " to ", upperlimit))
```

### Now do a CI for a median:

```{r medianCI}

os <- c(13, 12, 25, 17, 5)
PE <- median(os)

#  The code is really all the same, except changing means to medians in the PE and the bootstrap samples.
Allmedians <- c()
for (i in 1:10000) {
  s <- sample(os, size = length(os), replace = TRUE)
  Allmedians[i] <- median(s)
}

#  Allmeans
hist(Allmedians)

SE <- sd(Allmedians)
SE
ME <- 2 * SE

lowerlimit <- PE - ME
upperlimit <- PE + ME

lowerlimit <- round(lowerlimit, 2)
upperlimit <- round(upperlimit, 2)

print(paste("I'm 95% confident that the true median number of states visited by students in 087 is captured in the interval from ", lowerlimit, " to ", upperlimit))


```


###  Now find a CI for theWetsuit example

#### Note how to deal with a missing value!

```{r wetsuits}
os <- c(0.08,0.10,0.07,0.08,0.10,0.11,
        0.05,0.05,0.06,0.08,NA,0.05,0.10)

os <- na.omit(os)    # to remove missing values from vector

PE <- mean(os)
PE
#  CI for a mean

Allmeans <- c()
for (i in 1:10000) {
  s <- sample(os, size = length(os), replace = TRUE)
  Allmeans[i] <- mean(s)
}

#  Allmeans
hist(Allmeans)

SE <- sd(Allmeans)
SE

ME <- 2 * SE

lowerlimit <- PE - ME
upperlimit <- PE + ME

lowerlimit <- round(lowerlimit, 3)
upperlimit <- round(upperlimit, 3)

print(paste("I'm 95% confident that the true mean difference in max swimming velocity wearing a wetsuit compared to not wearing one is captured in the interval from ", lowerlimit, " to ", upperlimit))

```

###  Find a CI for the mean REM sleep time of animals

###  We'll use a vector in the data frame, msleep

```{r sleep, message = FALSE}

library(tidyverse)
tibble(msleep)

os <- msleep$sleep_rem
os <- na.omit(os)
  
  
PE <- mean(os)
PE

#  CI for a mean

Allmeans <- c()
for (i in 1:10000) {
  s <- sample(os, size = length(os), replace = TRUE)
  Allmeans[i] <- mean(s)
}

#  Allmeans
hist(Allmeans)

SE <- sd(Allmeans)
SE

ME <- 2 * SE

lowerlimit <- PE - ME
upperlimit <- PE + ME

lowerlimit <- round(lowerlimit, 3)
upperlimit <- round(upperlimit, 3)

print(paste("I'm 95% confident that the true mean REM sleep hours for animals is captured in the interval from ", lowerlimit, " to ", upperlimit))

```



