---
title: "Kissing the Right Way"
author: "Syourname"
date: "11/9/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

### Making up Sample Statistics

Make up 10 possible outcomes of the study, and measure the variation. These are just guesses to get you thinking about how statistics might vary from study to study.

```{r madeupstats}
#   Create a vector with 10 possible outcomes of percentage kissing right for 124 couples:


# Find the standard deviation of the  vector.


#  This value is the STANDARD ERROR, because it is the standard deviation of a bunch of statistics.  Statistics are values that summarize studies.  Here, we are looking at statistics that are percentages.
```

### Now simulate outcomes of the study in a better way:  Bootstrap Samples

For percentages, we use rbinom() to create bootstrap samples.

```{r bootstrap}
#  To simulate samples in a better way, "redo the study"  by randomly generating 124 couples kissing, with a probability of 80/124 of kissing right.
#  Trials will be the number of couples who kiss right.


#  Do many Trials by increasing n.  Just print the first few values.


#  This creates the NUMBER of couples kissing Right,
#  So make it a proportion.  Just print the first few values.


#  Do table() and hist()

```

### Use the Standard Error to estimate the precision of the original study.

```{r SE}
#  The standard deviation of the Bootsrap Kiss Percentages (the SE) will help us assess the precision of the study of 124 couples.

#  Two times the SE gives you the Margin of Error -- the ME


```

### Now, make a 95% Confidence Interval

```{r CI}
# The point estimate is the original percentage from the original data.

#  The CI is the PE plus or minus the ME


# print(paste("I'm 95% confident the true percentage of all couples who kiss on the right is captured in the interval from ", lowerlimit, " to ",upperlimit, "%"))
```

###  Vermonter poll example:

```{r Vpoll}
#  Simulate many bootstrap samples

#  This creates the NUMBER of people saying sty in VT,
#  So make it a proportion:


#  Calculate the Standard Error


#  Then the Margin of Error


#  The point estimate was given to be 47

#  State the CI:

```

