---
title: "Bechdel Test"
author: "yourname"
date: "10/19/2021"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

b <- read.csv("bechdel87.csv")
tibble(b)

```

### Graphs of categorical variables:

####  (Decade is also numeric, but we're treating it like a categorical variable just for this graph)

```{r categorical}
ggplot(data = b,
       mapping = aes(x = Decade, color = Decade)) + 
  geom_bar()

ggplot(data = b,
       mapping = aes(x = Btest, fill = Btest)) + 
  geom_bar()

#  Repeat for Btest5:
ggplot(data = b,
       mapping = aes(x= Btest5, fill = Btest5)) +
  geom_bar()

#  To find percentage:
t <- table(b$Btest)
prop.table(t)

```

###  Quantitative variables

####  Film budget (in 2013 dollars)

```{r budget, message = FALSE, warning = FALSE}
ggplot(data = b,
       mapping = aes(x = budget13)) +
  geom_histogram(fill = 'darkgreen', color = 'white')
```

#### Domestic gross income from film (just money made in U.S.) (in 2013 dollars)

```{r domgross, message = FALSE, warning = FALSE}
ggplot(data = b,
       mapping = aes(x = domgross13)) +
  geom_histogram(fill = 'darkgreen', color = 'white')
```


#### International gross income (includes domestic income, too) (in 2013 dollars)
# very skewed to the right

```{r intgross, message = FALSE, warning = FALSE}
ggplot(data = b,
       mapping = aes(x = intgross13)) + # inflation is ...
  geom_histogram(fill = 'darkgreen', color = 'white')
```

#### Crowd-sourced movie ratings.
#### Scale is from 1 (worst) to 5 (best)
# skewed to the left

```{r ratings, message = FALSE, warning = FALSE}

```

### Relationships among variables:

#### Bechdel Test by Decade

```{r btestdecade}
ggplot(data = b, 
       mapping = aes(x=Btest, color = Btest)) +
  geom_density() +
  facet_grid(~averating)
```

#### Average Rating by whether film passed the test:

```{r ratingbybt, messsage = FALSE, warning = FALSE}
ggplot(data = b, 
       mapping = aes(x=averating, fill = Btest)) +
  geom_density() +
  facet_wrap(~Btest, ncol=1)

```

#### Budget by whether film passed the test:

```{r budgbybt, message = FALSE, warning = FALSE}


```

#### Domestic Gross by whether film passed the test:


```{r domgrossbybt, message = FALSE, warning = FALSE}

```

####  Relationship between budget and gross earnings?

```{r budgetearn, warning = FALSE, message = FALSE}

```

### Create a new variable that is the percentage profit:

```{r percprofit}
# b2 <- b %>% 
#   mutate(domprofit13 = 
#            100*(domgross13 - budget13)/budget13)
```

### Percent Profit by Btest:
####  Better to just do the median, since median is better measure of Center when data is really skewed:

```{r ppbybt}

```

####  Conclusion: 
