---
title: "dplyr basics"
author: "Sheila Weaver"
date: "9/28/2021"
output: html_document
---

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#  Load the tidyverse
library(tidyverse)

#  Or, just load dplyr and ggplot2...
```

###  dplyr for data transformation
### 5MV = Five Main Verbs

####  1.  **filter** -- choosing rows

```{r filter}
#  Use the mpg data and keep only the cars with hwy mileage
#  better than 40
# filter keeps all of the rows that make the logical expression
#  in the parentheses TRUE

mpg40 <- mpg %>% filter(hwy > 40)
tibble(mpg40)

#   keep cars with hwy > 40  AND cty > 30
#  either of these works:
mpg3040 <- mpg %>% filter(hwy > 40 & cty > 30)
tibble(mpg3040)
mpg3040 <- mpg %>% filter(hwy > 40 , cty > 30)
tibble(mpg3040)

#  keep cars with hwy > 40 OR cty > 30
mpg3040 <- mpg %>% filter(hwy > 40 | cty > 30)
tibble(mpg3040)

#  keep cars that are suv's
suv <- mpg %>% filter(class == 'suv')
tibble(suv)

#  Keep all suvs and pickups
suvpu <- mpg %>% filter(class == 'suv' | class == 'pickup')
tibble(suvpu)
#  does the same thing:
suvpu <- mpg %>% filter(class %in% c('suv','pickup'))
tibble(suvpu)
```

#### filtering out missing values

```{r filtermiss}
#  here's something we use a lot:
#   filtering out the rows that have missing values for one variable:
#  If hwy had missing values:
mpgnomiss <- mpg %>% filter(!is.na(hwy))



```

####  2.  **mutate** -- use to create new variables and tranform old ones

```{r mutate}
#  Create a new variable, avemile, that is the average of city and highway mileage

mpg2 <- mpg %>% mutate(avemile = (cty + hwy)/2)


#  Confirm that it is in the data frame:
tibble(mpg2)

```

####  3.  **group_by** -- put cases in groups for analysis

```{r  group_by}
#  Does the new data frame look different?
#  NO!   need to tie group_by in with 'analysis'
#   See below....
mpgby <- mpg %>% group_by(class)
tibble(mpgby)

```

#### 4. **summarise** -- calculate stats on variables (columns)

```{r summarise}
#  Find mean, median, max and min of highway mileage:
mpg %>% summarise(meanhwy = mean(hwy),
                  medhwy = median(hwy),
                  maxhwy = max(hwy),
                  minhwy = min(hwy),
                  sampsize = n())


#  NOTE:  IF there were missing values, you'd need to something like this:
mpg %>% summarise(meanhwy = mean(hwy, na.rm = TRUE),
                  medhwy = median(hwy, na.rm = TRUE),
                  maxhwy = max(hwy, na.rm = TRUE),
                  minhwy = min(hwy, na.rm = TRUE),
                  sampsize = n())

#  OR... you could also use filter to remove rows missing hwy:
mpgnomiss <- mpg %>% filter(!is.na(hwy))
mpgnomiss %>% summarise(meanhwy = mean(hwy),
                  medhwy = median(hwy),
                  maxhwy = max(hwy),
                  minhwy = min(hwy),
                  sampsize = n())



#  Often useful to use with group_by.  Try redoing, grouping by class:
mpg %>% group_by(class) %>% 
                  summarise(meanhwy = mean(hwy),
                  medhwy = median(hwy),
                  maxhwy = max(hwy),
                  minhwy = min(hwy),
                  sampsize = n())

#  Try redoing, grouping by class and drv:
mpg %>% group_by(class, drv) %>% 
                  summarise(meanhwy = mean(hwy),
                  medhwy = median(hwy),
                  maxhwy = max(hwy),
                  minhwy = min(hwy),
                  sampsize = n())
```

####  5. **arrange** -- arranges rows by one or more variables

```{r}
#  arrange mpg data frame in order from lowest to highest highway mileage:
mpgarr <- mpg %>% arrange(hwy) 
tibble(mpgarr)

#  largest to smallest:
mpgarr <- mpg %>% arrange(desc(hwy))
tibble(mpgarr)

#  Top 10 cars for low mileage
mpg %>% arrange(desc(hwy)) %>% head(10)

#  Could arrange by more variables.  Categorical variables are arranged in alphabetical order.

```

### Other useful verbs:

####  **select** --  keep certain columns in the data

```{r select}
#  Keep only the variables manufacturer, class, hwy, cty:
mpg4 <- mpg %>% select(manufacturer, class, hwy, cty)
tibble(mpg4)

#  Keep all of the variables *except* hwy:
mpgnohwy <- mpg %>% select(-hwy)
tibble(mpgnohwy)

#  Using select might be useful for printing your data, or for saving a smaller version.

```

####  **sample_n** takes a random sample of rows in your data frame:

```{r sample_n}
#  print 4 randomly selected rows from the data:
mpg %>% sample_n(4)

#  Make a new data frame with only 20 rows:
mpg %>% sample_n(20)

#  One use of sample_n is when you have a huge data frame, and would like to try out graphs and analyses with a smaller set first, to keep down processing time.
```

#### Merging data frames with:   **left_join**, **right_join**, **inner_join**,  **full_join**

####  Save this for later....

