---
title: "Prediction Modeling"
author: "yourname"
date: "10/26/2021"
output: html_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#  load the tidyverse
library(tidyverse)

# read in the data and print as a tibble
c <- read.csv("Crickets.csv")

```

### Plot the data with a regression line

Describe the relationship here (Using points 1 through 4)

This is a strong, positive linear relationship with no outliers.   Crickets chirps increase with higher temperatures. 

```{r cricket, message = FALSE}
#  Use ggplot to graph the data, adding a regression line with geom_smooth(method = 'lm')
tibble(c)

ggplot(data = c,
      mapping = aes(x = Chirps, y = TempF)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)

```

### Find the regression line equation

Once you have the model, write the slope here in terms of the problem:  For each increase ....

```{r linearmodel}
# Create the linear model for temperature by chirps.  Be sure to state the y variable first -- y ~ x.  "Y By X"
CricketReg <- lm(c$TempF ~ c$Chirps)

CricketReg <- lm(TempF ~ Chirps, data = c)

#  Print the model object
CricketReg

#  Do names(CricketReg) to see what else is in the model object
names(CricketReg)
#  Do a summary of the model object
summary(CricketReg)
```

### Find the sample correlation coefficient

``` {r corr}

#  Print the correlation coefficient of temperature and chirps.  Try putting the temperature first and chirps second, then put in the reverse order:  

cor(c$Chirps, c$TempF)

cor(c$TempF, c$Chirps)

#  For the correlation coefficient, it does not matter what order you put the variables in.

#  If there were missing values in either of the variables, you could do this:

cor(c$TempF, c$Chirps, use = 'complete.obs')

```

###  Another example:  Using mtcars data, look at mpg by car weight

After you run the code below, describe the relationship here, using points 1 through 4.  Then state the slope in terms of the problem.

The relationship is strong and negative (r = -.867), and linear, with possible two outliers on the high mpg side, and a group of three very large vehicles with low mpg.

For each additional thousand pounds of car weight, there is a -5.3445 increase in predicted miles per gallon.  

For each additional thousand pounds of car weight, there is a 5.3445 decrease in predicted miles per gallon.   

```{r cardata, message = FALSE}
# Plot mpg by wt here, with a regression line.
ggplot(data = mtcars,
       mapping = aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)

# Create and summarize the model object
mo <- lm(mtcars$mpg ~ mtcars$wt)
summary(mo)

#  Find the correlation coefficient
cor(mtcars$mpg, mtcars$wt)
```

