---
title: "dplyr basics"
author: "Sheila Weaver"
date: "9/28/2021"
output: html_document
---

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#  Load the tidyverse
library(tidyverse)

#  Or, just load dplyr and ggplot2...
```

###  dplyr for data transformation
### 5MV = Five Main Verbs

####  1.  **filter** -- choosing rows

```{r filter}
tibble(mpg)

# Use the mpg data and keep only the cars with hmy mileage better than 40
mpg40 <- mpg %>% filter(hwy > 40)
tibble(mpg40)

# keep cars with hwy > 40 AND cty > 30 
# either commands work 
mpg3040 <- mpg %>% filter(hwy > 40 & cty > 30)
tibble(mpg3040)
mpg3040 <- mpg %>% filter(hwy > 40, cty > 30)
tibble(mpg3040)

# keep cars with hwy > 40 OR cty > 30 
mpg3040 <- mpg %>% filter(hwy > 40 | cty > 30)
tibble(mpg3040)

# keep cars that are suv's
mpgSuv <- mpg %>% filter(class == "suv")
tibble(mpgSuv)

# keep all suvs and pickups
suvpu <- mpg  %>% filter(class == "suv" |class == "pickup")
tibble(suvpu)
# does the same thing:
suvpu <- mpg  %>% filter(class %ini% c('suv','pickup'))
tibble(suvpu)
```

#### filtering out missing values

```{r filtermiss}
#  here's something we use a lot:
#   filtering out the rows that have missing values for one variable:
mpgnomiss <- mpg %>% filter(!is.no(hwy)) # takes away the lines that have missing hwy

```

####  2.  **mutate** -- use to create new variables and tranform old ones

```{r mutate}
#  Create a new variable, avemile, that is the average of city and highway mileage

#  Confirm that it is in the data frame:


```

####  3.  **group_by** -- put cases in groups for analysis

```{r  group_by}
#  Does the new data frame look different?


```

#### 4. **summarise** -- calculate stats on variables (columns)

```{r summarise}
#  Find mean, median, max and min of highway mileage:

#  NOTE:  IF there were missing values, you'd need to something like this:

#  OR... you could also use filter to remove rows missing hwy:

#  Back to the original -- now add the function n() to find the total number of observations.

#  Often useful to use with group_by.  Try redoing, grouping by class:

#  Try redoing, grouping by class and drv:
```

####  5. **arrange** -- arranges rows by one or more variables

```{r}
#  arrange mpg data frame in order from lowest to highest highway mileage:

#  largest to smallest:

#  Could arrange by more variables.  Categorical variables are arranged in alphabetical order.

```

### Other useful verbs:

####  **select** --  keep certain columns in the data

```{r select}
#  Keep only the variables manufacturer, class, hwy, cty:

#  Keep all of the variables *except* hwy:

#  Using select might be useful for printing your data, or for saving a smaller version.

```

####  **sample_n** takes a random sample of rows in your data frame:

```{r sample_n}
#  print 4 randomly selected rows from the data:

#  Make a new data frame with only 20 rows:

#  One use of sample_n is when you have a huge data frame, and would like to try out graphs and analyses with a smaller set first, to keep down processing time.
```

#### Merging data frames with:   **left_join**, **right_join**, **inner_join**,  **full_join**

####  Save this for later....

