---
title: "Practicing HTs and Bootstrap CIs"
author: "yourname"
date: "11/16/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

### Hypothesis Test

In his lifetime, Paul the Octopus guessed the outcomes of 12 out of 14 world cup soccer games correctly. 

Do a hypothesis test to see if he could have just done this by chance, or if there is something else at play...

```{r setupHT}
#   Calculate the observed result:


#  Write down the hypotheses:

#  Null Hypothesis:  
#  Alternative Hypothesis:  

```

#### Now simulate outcomes of the study under the null hypothesis

```{r modelHT}
#  Use rbinom to generate Trials
#  Trials <- rbinom(n = , size = , prob = )

#  Make a histogram of the trials

#  Make a table of the trials

```

#### Calculate the pvalue and make a decision

```{r pvalue}


#  This p-value represents the probabity that Paul got 12 (or more) correct guesses just by chance.  It is very  

#  Why did it happen?

```

###  Bootstrap Confidence Interval for a Percentage

We could also answer the same question with a confidence interval.Use rbinom() to create bootstrap samples.

```{r bootstrap}
#  Calculate the PE

#  Generate trials using the PE for the probability
# Trials <- rbinom(n = , size = , prob = )

#  Express the trials as percentages;
# Perc <- 100 * Trials/14

#  Do  hist()

```

#### Use the Standard Error to estimate the precision of the original study.

```{r SE}
#  The standard deviation of the Bootstrap percentages is the SE


#  Two times the SE gives you the Margin of Error -- the ME


```

#### Now, make a 95% Confidence Interval

```{r CI}
# The point estimate is the original percentage from the original data.

#  The CI is the PE plus or minus the ME
# upperlimit <- PE + ME
# lowerlimit <- PE - ME
# 
# upperlimit <- round(upperlimit, 2)
# lowerlimit <- round(lowerlimit, 2)
# 
# print(paste("I'm 95% confident the true accuracy rate of Paul the Octopus, if he could keep predicting  is captured in the interval from ", lowerlimit, " to ",upperlimit, "%"))

#  Is he doing better than chance?
```

### Bootstrap Confidence Interval for a Mean 

Find a confidence interval for true mean REM sleep for animals.  Assume msleep contains a random sample of animals, and use the variable sleep_rem

```{r sleeprem}

#  Summarise the vector:

#  Let it equal os

#  Remove missing values from os

#  Find the point estimate -- the mean of os

#  Make many bootstrap samples from os, and save in the vector Allmeans:


#  Do a histogram of Allmeans

#  Find the SE, the ME, and the confidence interval


#  State the CI:
# print(paste("I'm 95% sure that the true mean rem sleep time for animals is captured in the interval from ", lowerlimit, " to ", upperlimit, " hours."))

```

