---
title: "Modeling Random Behavior"
author: "yourname"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Modeling Coin Flips

#### One coin:

```{r onecoin}
#  Having R flip one coin:
Trial <- rbinom(n = 1, size = 1, prob = .5)
ifelse(Trial == 1, 'Heads', 'Tails')
```

#### 10 coins at once, Heads and Tails

``` {r tencoins}
Trial <- rbinom(n = 10, size = 1, prob = .5)
ifelse(Trial == 1, 'Heads', 'Tails')
```

#### 10 coins at once, printing Trial Result:  number of heads

```{r trialresult}
#  To do sets of 10 -- this prints the 
#  "Trial Result"  -- the number of
#   heads in one trial of 10 tosses:
Trial <- rbinom(n = 1, size = 10, prob = .5)
Trial
```

#### Do a bunch of trials of 10 coins

```{r manytrials}
#  Same thing with bigger n
Trial <- rbinom(n = 100, size = 10, prob = .5)
Trial
hist(Trial)
table(Trial)
```

### Modeling Free Throws for a 90% shooter

```{r freethrows}
#  Having R do one free throw:
Trial <- rbinom(n = 1, size = 1, prob = .9)
ifelse(Trial == 1, 'Made', 'Miss')

#  To do all 10 at once:
Trial <- rbinom(n = 10, size = 1, prob = .9)
ifelse(Trial == 1, 'Made', 'Miss')

#  To do sets of 10 -- this prints the 
#  "Trial Result"  -- the number of
#   Made free throws in 10 attempts: 
Trial <- rbinom(n = 1, size = 10, prob = .9)
Trial 

#    Do a bunch of trials:
Trial <- rbinom(n = 100000, size = 10, prob = .9)
# Trial 
hist(Trial)
table(Trial)
```


### To do the Burning House hypothesis test
#### Simulate 17 choices of house, if houses are equally likely

```{r burninghouse}
14/17

Trials <- rbinom(n = 1000, size = 17, prob = .5)
hist(Trials)
table(Trials)

#  How often did values of 14 or higher come up just by chance
#  For me, this happened 11/1000 = .0011

#  So she could have chosen 14 (or more) just by chance.
#  But the chance of this is really low (.0011) 
#  (and a common cutoff for "LOW" is < .05)
#  So we reject this idea, and conclude that the 
#  alternative hypothesis is true:  SHE DOES have a preference

```

###  Lady Tasting Tea -- got all 8 correct -- milk or tea first

```{r tea}
#  Null:  She can't tell the difference 
#  Alternative:  She can

Trials <- rbinom(n = 1000, size = 8, prob = .5)
hist(Trials)
table(Trials)

#  She could have gotten all 8 correct by chance,
#  but it only happened 6/1000 times for me .006

#  Conclusion:  This is so unlikely that I conclude that 
#   she really knows the difference.

```


### To do the 80% shooter hypothesis test
#### Simulate 20 free throws, if person is really an 80% shooter

```{r}
#  Observed result = 13/20 = .65


# Null Hypothesis:  Really is an 80% shooter
# Alternative Hypothesis:   Really is worse than 80%

Trials <- rbinom(n = 10000, size = 20, prob = .8)
hist(Trials)
table(Trials)

#  Getting 13 or worse could happen by chance.  
#  in my simulation, it happened 91 times out of 1000
#  so  .091  <-- this is called the "p value"
 5+26+60

#  This is not small enough to reject the Null, so we 
#  conclude that we do NOT have enough evidence to say 
#  that he is worse than an 80% shooter.
```

