---
title: "HW#3"
author: "Lucía Carrera"
date: "10/1/2021"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#  load the tidyverse here -- OR-- load dplyr and ggplot2
library(tidyverse)

```

##  Question 1: Mapping Aesthetics -- 4 pts

####  Variable -- Aesthetic 

- Year -> x
- Totalprop -> y
- Letter -> facet
- Sex -> fill

<put your ggplot code here>
ggplot(data = babies, mapping = aes(x=year, y=totalprop, color=sex)) +
  geom_line()+
  facet_wrap(~ letter)


##  Question 2:  dplyr 

### part a  -- lists in order  3 pts

```{r top5}
# View and Help
# View(msleep)
# help(msleep)

# Method 1
msleep %>% filter(!is.na(awake),!is.na(name)) %>%  arrange(desc(awake)) %>% select(name,awake) %>% head(6)

# Method 2
# msleep_nameAwake <- msleep %>% select(name,awake) 
# msleep_nameAwake %>% arrange(desc(awake)) %>% head(6)

```

```{r last5}
msleep %>% filter(!is.na(awake),!is.na(name)) %>% arrange(awake) %>% select(name,awake) %>% head(6)

```

### part b -- group_by -- 3 pts

```{r groupby}
mpgby <- msleep %>% group_by(sleep_rem)
msleep %>% group_by(vore) %>% filter(!is.na(vore)) %>% 
                  summarise(meanawake = mean(awake),
                  sdawake = sd(awake),
                  n_vore = n())
                  
```

##  Question 3 

### part a -- bar graph -- 3 pts

```{r bargraph}

vores_nona <- msleep %>% filter(!is.na(conservation), !is.na(vore))

ggplot(data = vores_nona,
       mapping = aes( x = vore, fill = conservation)) +
    geom_bar(position = 'fill') +
    labs(subtitle = 'Conservation status of animal by "vore"')
```

### part b -- box plot -- 3 pts

```{r boxplot}
msleep %>% filter(!is.na(vore)) %>%
  ggplot(mapping = aes(x = vore, y = sleep_total, fill = vore)) +
  geom_boxplot() +
  labs( title = 'Total Hours of Sleep by "vore"', y = 'Total Hours of Sleep')

```



### part c -- facets -- 3 pts

```{r facets}
msleep %>% filter(!is.na(vore)) %>%
  ggplot(mapping = aes(x = sleep_total)) +
  geom_density() +
  labs( title = 'Total Hours of Sleep by "vore"', x = 'Total Hours of Sleep') +
  facet_wrap(~ vore, ncol = 1)
```

  
### part d -- logtrans -- 3 pts

```{r logtrans, message = FALSE, warning = FALSE}
msleep %>% filter(!is.na(vore)) %>%
  ggplot(mapping = aes(x = bodywt, y=brainwt)) +
  geom_point(size=.9) +
  geom_smooth() +
  scale_x_log10() +  
  scale_y_log10() +
  labs( title = 'Brain Weight by Body Weight (log scale)', x = 'Body Weight (log scale)', y = 'Brain Weight (log scale)') +
  theme_light()

```

### part e -- whylog -- 4 pts

<why is the log transformation useful?  Put answer here >
The log transformation is used when our original data does not follow the bell curve. With this mathematical tool we can transform the data so that its base is not as skewed. If we observe the graph below, most of the points are in the left and then the gray region occupies a large part of the graph. This is due to the data being skewed.

```{r nolog, warning = FALSE}
#  Graph here the same scatterplot without the log transformation
msleep %>% filter(!is.na(vore)) %>%
  ggplot(mapping = aes(x = bodywt, y=brainwt)) +
  geom_point(size=.9) +
  geom_smooth() +
  labs( title = 'Brain Weight by Body Weight (log scale)', x = 'Body Weight (log scale)', y = 'Brain Weight (log scale)') +
  theme_light()
```



##  Question 4:   Lie Factor --   4 pts 

< Type in calculations, answer, and describe the result here >
Using the following [website] (https://eleif.net/photo_measure.html) I was able to measure the exact graphical value of the bars. 
Effect in...


<table>
<th>
</th>
<th>
Graphic             
</th>
<th>
Data                
</th>
</tr>
<td>
Barry Bonds bar     
</td>
<td>
7.2                 
</td>
<td>
762
</td>
</tr>
<tr>
<td>
Babe Ruth bar
</td>
<td>
2.4           
</td>
<td>
714
</td>
</tr>
</table>



#### Formula
LIE_effect = SIZE_effect_graphic / SIZE_effect_data
SIZE = ( VALUE_one - VALUE_two ) / VALUE_two


#### Comparison (Barry Bonds and Babe Ruth bar):
SIZE_effect_graphic = (( 7.2 - 2.4 ) / 2.4 ) * 100 = 200%
SIZE_effect_data = (( 762 - 714 ) / 714 ) * 100 = 6.72%

LIE_effect = 200 / 6.72 = 29.76


#### Conclusion
The Lie effect of this graph is not in range according to Tufte (0.95 to 1.05). In fact, it is very much over the desired number, meaning that the graph presented overstates the effect in the data. The bargraph is not directly proportional to the numerical data, hence we should not be making conclusions of the data collected according to it. Without making calculations, one could see that there is a potential problem by seeing that it does not commence in zero but instead at 690. If the entire y axis had been shown we would have seen that there is very little difference between each bar.

